---
title: "Regular Expressions"
subtitle: "ISAAK short lesson"
date: 2018-10-24
author: "Clemens Schmid"
output: beamer_presentation
editor_options: 
  chunk_output_type: console
fontsize: 9pt
---

## Motivation

```{r}
last_names <- c("Maier", "Meyer", "Tietze", "Mayr", "Rinne")
```

```{r, echo = FALSE}
last_names
```

```{r}
grep(pattern = "M[ae][iy]e?r", last_names)

last_names[grep(pattern = "M[ae][iy]e?r", last_names)]
```

## Definition

- Regular Expressions are sign sequences that define specific search patterns
- Standards: 
  - POSIX: Basic Regular Syntax (BRE)
  - POSIX: Extended Regular Syntax (ERE) 
  - Perl regexes

## Symbols

|Expression|Meaning|Example|
|---|---|---|
|c|Individual signs|c|
|.|One sign, except linebreak|.|
|[sign]|One of these signs|[123ufg]|
|[sign1 - sign2]|One sign from this sequence|[0-9]|
|[^sign]|One sign that is NOT one of these|[^a]|
|[^sign1 - sign2]|One sign NOT from this sequence|[^0-9]|

## Quantifiers

|Expression|Meaning|
|---|---|
|?|sign before ? zero times or once|
|$\ast$|sign before $\ast$ zero times or or any number of times|
|+|sign before + once or any number of times|
|{n}|sign before {n} n times|
|{n,m}|sign before {n,m} n to m times|
|{n,}|sign before {n,} minimum n times|
|{,m}|sign before {,m} maximum n times|

## Other Expressions

|Expression|Meaning|
|---|---|
|$\widehat{}$|Start of the line|
|$|End of line|
|\\<|Start of word|
|\\<|End of word|
|(...)|Subexpression. Important for replacement commands|
|\||Logical OR|
|\\|Avoid special meaning of operators after \\|



