---
title: "Regular Expressions"
subtitle: "ISAAK short lesson"
date: 2018-10-24
author: "Clemens Schmid"
output: beamer_presentation
editor_options: 
  chunk_output_type: console
fontsize: 9pt
---

## Motivation

```{r}
last_names <- c("Maier", "Meyer", "Tietze", "Mayr", "Rinne")
```

```{r, echo = FALSE}
last_names
```

```{r}
grep(pattern = "M[ae][iy]e?r", last_names)

last_names[grep(pattern = "M[ae][iy]e?r", last_names)]
```

## Definition

- Regular Expressions are character sequences that define specific search patterns
- Standards: 
  - POSIX: Basic Regular Syntax (BRE)
  - POSIX: Extended Regular Syntax (ERE) 
  - Perl regexes

## Symbols

|Expression|Meaning|Example|
|---|---|---|
|c|Individual characters|c|
|.|One character, except linebreak|.|
|[character]|One of these characters|[123ufg]|
|[character1 - character2]|One character from this sequence|[0-9]|
|[^character]|One character that is NOT one of these|[^a]|
|[^character1 - character2]|One character NOT from this sequence|[^0-9]|

