---
title: "Regular Expressions in R"
subtitle: "ISAAK short lesson"
date: 2018-10-24
author: "Clemens Schmid"
output: beamer_presentation
editor_options: 
  chunk_output_type: console
fontsize: 9pt
---

## Motivation

```{r}
last_names <- c("Maier", "Meyer", "Tietze", "Mayr", "Rinne")
```

```{r, echo = FALSE}
last_names
```

```{r}
grep(pattern = "M[ae][iy]e?r", last_names)

last_names[grep(pattern = "M[ae][iy]e?r", last_names)]
```

# Regular Expressions

## Definition

- Regular Expressions are sign sequences that define specific search patterns
- Standards: 
  - POSIX: Basic Regular Syntax (BRE)
  - POSIX: Extended Regular Syntax (ERE) 
  - Perl regexes

## Symbols

|Expression|Meaning|Example|
|---|---|---|
|c|Individual signs|c|
|.|One sign, except linebreak|.|
|[sign]|One of these signs|[123ufg]|
|[sign1 - sign2]|One sign from this sequence|[0-9]|
|[^sign]|One sign that is NOT one of these|[^a]|
|[^sign1 - sign2]|One sign NOT from this sequence|[^0-9]|

## Quantifiers

|Expression|Meaning|
|---|---|
|?|sign before ? zero times or once|
|$\ast$|sign before $\ast$ zero times or or any number of times|
|+|sign before + once or any number of times|
|{n}|sign before {n} n times|
|{n,m}|sign before {n,m} n to m times|
|{n,}|sign before {n,} minimum n times|
|{,m}|sign before {,m} maximum n times|

## Other Expressions

|Expression|Meaning|
|---|---|
|$\widehat{}$|Start of the line|
|$|End of line|
|\\<|Start of word|
|\\<|End of word|
|(...)|Subexpression. Important for replacement commands|
|\||Logical OR|
|\\|Avoid special meaning of operators after \\|

# Regular Expressions in R

## Pattern Matching and Replacement in base R (?grep)

```{r, eval = FALSE}
grep(pattern, x, ignore.case = FALSE, perl = FALSE, value = FALSE,
     fixed = FALSE, useBytes = FALSE, invert = FALSE)

grepl(pattern, x, ignore.case = FALSE, perl = FALSE,
      fixed = FALSE, useBytes = FALSE)

sub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE,
    fixed = FALSE, useBytes = FALSE)

gsub(pattern, replacement, x, ignore.case = FALSE, perl = FALSE,
     fixed = FALSE, useBytes = FALSE)

regexpr(pattern, text, ignore.case = FALSE, perl = FALSE,
        fixed = FALSE, useBytes = FALSE)

gregexpr(pattern, text, ignore.case = FALSE, perl = FALSE,
         fixed = FALSE, useBytes = FALSE)

regexec(pattern, text, ignore.case = FALSE, perl = FALSE,
        fixed = FALSE, useBytes = FALSE)
```

## Pattern Matching and Replacement in base R (?grep)

- **grep** & **grepl**: Search a character vector and return the indices of the matching elements. **grep** returns a vector of the indices. **grepl** returns a logical vector (TRUE/FALSE).

- **sub** & **gsub**: Replace. Search a character vector for regular expression matches and replace that match with another string. **sub** replaces the first occurrence of a pattern, **gsub** replaces all occurrences.

- **regexpr** & **gregexpr** & **regexec**: Search a character vector for matches and return the indices of the string where the match begins and the length of the match. **gregexpr** returns result as a list object. **regexec** additionally returns the locations of any parenthesized sub-expressions. 

# Example Code

## Get Data

```{r}
library(magrittr)

radon <- c14bazAAR::get_RADON()[,c("labnr", "shortref")] %>% 
  tibble::as.tibble()
```

```{r, echo = FALSE}
radon
```

## Search for "Breunig 1987"

```{r}
radon %>%
  dplyr::filter(
    shortref == "Breunig 1987"
  )
```

## Search for "Breunig 1987"

```{r}
radon %>%
  dplyr::filter(
    grepl("Breunig 1987", .$shortref)
  ) 
```

## Search for "Breunig 1987" from page 64-78 

```{r}
radon %>%
  dplyr::filter(
    grepl("Breunig 1987,?\\s+[64-78]", .$shortref)
  ) 
```

## Search for "Breunig 1987" from page 64 without upper limit

```{r}
radon %>%
  dplyr::filter(
    grepl("Breunig 1987,?\\s+([6-9]|[1-9]{3,})", .$shortref)
  ) 
```
